### 2022-06-27 main/chapter_2ljets.tex
some spelling grammar check

That is, finely binning discovery regions can harm sensitivity since their
interpretations are decoupled.
Since we do couple the background models in exclusion studies,
opposite is true (to some extent) there.

@@

So far, we have seen many histograms of background samples overlaid with
observed data.
Those backgrounds may or may not have appeared to predicted the data well, but
no conclusions can be drawn until uncertainties are assigned to the background
and signal modelling.

@@

With such a complicated experiment as \atlas, in a universe with such varied
particles as ours, there are many different factors of uncertainty to consider.
As described in Section~\ref{sec:searches_practice}, the effect of each
factor is studied to assign regularized variations to the Poisson expectations
of the histogram bins with a HistFactory model~\cite{cranmer2012histfactory}.
For the $\twoljets$ search, we used HistFitter to build our every HistFactory
models~\cite{Besjes_2015,baak2015histfitter}.

@@

The $\twoljets$-electroweak search implements a total of $136$ modifiers its
background-only model, and more for each signal model.
Of those I evaluated more manually, some of the more interesting or important
will be discussed in the remainder of this section ---
Fake/non-prompt lepton backgrounds in Section~\ref{sec:2ljets_mm_fakes},
alternative simulated samples in Section~\ref{sec:2ljets_sample_comparisons},
signal modelling uncertainties in Section~\ref{sec:2ljets_signal_variations}
and theoretical uncertainties in Section~\ref{sec:2ljets_theory_variations}.

@@

Given our requirements on lepton quality, isolation, and hardness, it is
expected that the rates of fake leptons, or non-prompt leptons (which do not
originating from the hard scatter) should be small.
And indeed more details estimates indicate that they are small.
However, just like other backgrounds, it is important that we assign
honest and honestly uncertain estimates for these fake/non-prompt backgrounds.

Although my style of spelling out fake/non-prompt every time
(as opposed to just `fake') may be awkward, it is supported by fact that this
the non-prompt part dominates.
Indeed, figures in our internal documentation~\cite{twoljets2018int} suggest
that, in simulations, non-prompt hadron decays produce in excess of $90\%$ of
this background in a loose selections near to our regions, and across the
entire $\metsig$ spectrum.
Non-prompt muons are mostly from decays of heavy-flavour ($b$ or $c$) hadrons,
and electrons arise in roughly equal parts from light flavour hadrons.
Tau decays form the next largest part, with prompt fakes due to misidentified
photons or leptons in the sub-$\%$ weeds.

@@

In data, one measures tightly ($t$) and loosely ($l$) identified leptons, leading
to four categories of events with two ordered leptons as labelled $tt$, $tl$,
$lt$, and $ll$ for tight-tight, tight-loose, and so on.
The Matrix Method assumes a linear relationship between these selections
and those of real ($r$) and fake/non-prompt ($f$) leptons:
\begin{equation}
\label{eqn:2ljets_matrix_method}
\begin{pmatrix}
\mu^\textrm{data}_{tt} \\
\mu^\textrm{data}_{tl} \\
\mu^\textrm{data}_{tl} \\
\mu^\textrm{data}_{ll} \\
\end{pmatrix}
=
M
\begin{pmatrix}
\mu^\textrm{truth}_{rr} \\
\mu^\textrm{truth}_{fr} \\
\mu^\textrm{truth}_{rf} \\
\mu^\textrm{truth}_{ff} \\
\end{pmatrix}
,
\end{equation}
where $\mu^{\ldots}_{**}$ is a Poisson event rate and $M$ is the matrix of
transition probabilities after which this method is
named~\cite{ATLAS-CONF-2014-058}.

@@

In fact, real data tend to be assigned negative Matrix Method weights.
For example, CR-VZ receives $157$ data in its Matrix Method sample,%
\footnote{%
This count of $157$ differs from the $194$ actual data observed in CR-VZ.
I do not know why.
The two do overlap in event identity numbers, so there is presumably some
tighter selection in the Matrix Method which should be accounted for in its
weights.
}
of which $135$ pass the signal lepton requirements (tight-tight),
and the remainder have one loose lepton (zero loose-loose lepton events
are observed in CR-VZ).
The tight-tight events have an average weight of $-0.0227$, and the others
have an average weight of $+0.238$ (one can check these recover the central
fake/non-prompt background of $2.2$).
Such negative signal weights suggest that a local \emph{excess} of data
would act to \emph{decrease} the Matrix Method estimate in the region.
Yes this backwards dependency is backward, but at least it gives some relief
hide real local excesses would not be hidden by increased fake estimates.

@@

\begin{table}[tp]
\centering
\begin{tabular}{lccc}
& MC Prompt (stat) & MC FNP (stat) & MM FNP (stat, sys)
\\[0.5em]
CR-Z & $143.6 \pm 4.6$ & $0.1 \pm 0.1$ & $-0.2 \pm 1.2 \pm 2.0$
\\
VR-Int & $18.7 \pm 0.9$ & $0.2 \pm 0.2$ & $-0.3 \pm 0.1 \pm 0.3$
\\
VR-High & $19.7 \pm 0.7$ & $0.04 \pm 0.04$ & $-0.2 \pm 0.2 \pm 0.5$
\\
VR-High-1J & $51.9 \pm 1.9$ & $0.003 \pm 0.002$ & $-1.9 \pm 0.2 \pm 0.4$
\\
SR-Low-b & $15.4 \pm 1.0$ & $0.001 \pm 0.000$ & $-0.08 \pm 0.05 \pm 0.06$
\\
SR-Int-a & $22.3 \pm 0.8$ & $0 \pm 0$ & $-0.4 \pm 0.2 \pm 0.6$
\\
SR-Int-b & $11.4 \pm 0.5$ & $0.002 \pm 0.002$ & $-0.3 \pm 0.1 \pm 0.3$
\\
SR-High-8-a & $2.4 \pm 0.2$ & $0 \pm 0$ & $0 \pm 0 \pm 0$
\\
SR-High-8-b & $2.4 \pm 0.3$ & $0 \pm 0$ & $0 \pm 0 \pm 0$
\\
SR-High-1J & $1.0 \pm 0.1$ & $0.004 \pm 0.003$ & $0 \pm 0 \pm 0$
\\
SR-$\llbb$ & $0.7 \pm 0.1$ & $0 \pm 0$ & $0 \pm 0 \pm 0$
\\
\end{tabular}
\caption[%
Background yields from prompt and fake/non-prompt sources in bins with
non-positive Matrix Method estimates
]{%
Background yields from prompt and fake/non-prompt sources in bins with
non-positive Matrix Method estimates.
The `MC FNP' yields are selected using truth information in sumulated samples.
The `MM FNP' yields are from the Matrix Method, with statistical errors from
the sample weights and systematic variations from the larger absolute change
from variations of those weights.
}
\label{tab:2ljets_fnp_negative_mm}
\end{table}

>>>>

That is, finely binning discovery regions can harm sensitivity since their
interpretations are decoupled.
Since we do couple the background models in exclusion studies,
the opposite is true (to some extent) there.

@@

So far, we have seen many histograms of background samples overlaid with
observed data.
Those backgrounds may or may not have appeared to have predicted the data well,
but no conclusions can be drawn until uncertainties are assigned to the
background and signal modelling.

@@

With such a complicated experiment as \atlas, in a universe with such varied
particles as ours, there are many and varied factors of uncertainty to consider.
As described in Section~\ref{sec:searches_practice}, the effect of each
factor is studied to assign regularized variations to the Poisson expectations
of the histogram bins with a HistFactory model~\cite{cranmer2012histfactory}.
For the $\twoljets$ search, we used HistFitter to build our every HistFactory
models~\cite{Besjes_2015,baak2015histfitter}.

@@

The $\twoljets$-electroweak search implements a total of $136$ modifiers in its
background-only model, and more for each signal model.
Of those I evaluated more manually, some of the more interesting or important
will be discussed in the remainder of this section ---
Fake/non-prompt lepton backgrounds in Section~\ref{sec:2ljets_mm_fakes},
alternative simulated samples in Section~\ref{sec:2ljets_sample_comparisons},
signal modelling uncertainties in Section~\ref{sec:2ljets_signal_variations}
and theoretical uncertainties in Section~\ref{sec:2ljets_theory_variations}.

@@

Given our requirements on lepton quality, isolation, and hardness, it is
expected that the rates of fake leptons, or non-prompt leptons (which do not
originating from the hard scatter) should be small.
Indeed, more detailed estimates indicate that they are small.
However, just like other backgrounds, it is important that we assign
honest and honestly uncertain estimates for these fake/non-prompt backgrounds.

Although my style of spelling out fake/non-prompt every time
(as opposed to just `fake') may be awkward, it is supported by the fact that this
the non-prompt part dominates.
Indeed, figures in our internal documentation~\cite{twoljets2018int} suggest
that, in simulations, non-prompt hadron decays produce in excess of $90\%$ of
this background in a loose selection near to our regions, and across the
entire $\metsig$ spectrum.
Non-prompt muons are mostly from decays of heavy-flavour ($b$ or $c$) hadrons,
and electrons arise in roughly equal parts from light flavour hadrons.
Tau decays form the next largest part, with prompt fakes due to misidentified
photons or leptons in the sub-$\%$ weeds.

@@

In data, one measures tightly ($t$) and loosely ($l$) identified leptons, leading
to four categories of events with two ordered leptons as labelled $tt$, $tl$,
$lt$, and $ll$ for tight-tight, tight-loose, and so on.
The Matrix Method assumes a linear relationship between these selections
and those of real ($r$) and fake/non-prompt ($f$) leptons:
\begin{equation}
\label{eqn:2ljets_matrix_method}
\begin{pmatrix}
\mu^\textrm{data}_{tt} \\
\mu^\textrm{data}_{tl} \\
\mu^\textrm{data}_{tl} \\
\mu^\textrm{data}_{ll} \\
\end{pmatrix}
=
M
\begin{pmatrix}
\mu^\textrm{truth}_{rr} \\
\mu^\textrm{truth}_{fr} \\
\mu^\textrm{truth}_{rf} \\
\mu^\textrm{truth}_{ff} \\
\end{pmatrix}
,
\end{equation}
where $\mu^{\ldots}_{**}$ is a Poisson event rate and $M$ is the matrix of
transition probabilities, after which this method is
named~\cite{ATLAS-CONF-2014-058}.

@@

In fact, real data tend to be assigned negative Matrix Method weights.
For example, CR-VZ receives $157$ data in its Matrix Method sample,%
\footnote{%
This count of $157$ differs from the $194$ actual data observed in CR-VZ.
I do not know why.
The two do overlap in event identity numbers, so there is presumably some
tighter selection in the Matrix Method, which should be accounted for in its
weights.
}
of which $135$ pass the signal lepton requirements (tight-tight),
and the remainder have one loose lepton (zero loose-loose lepton events
are observed in CR-VZ).
The tight-tight events have an average weight of $-0.0227$, and the others
have an average weight of $+0.238$ (one can check these recover the central
fake/non-prompt background of $2.2$).
Such negative signal weights suggest that a local \emph{excess} of data
would act to \emph{decrease} the Matrix Method estimate in the region.
Yes this backwards dependency is backward, but at least it gives some relief
hide real local excesses would not be hidden by increased fake estimates.

@@

\begin{table}[tp]
\centering
\begin{tabular}{lccc}
& MC Prompt (stat) & MC FNP (stat) & MM FNP (stat, sys)
\\[0.5em]
CR-Z & $143.6 \pm 4.6$ & $0.1 \pm 0.1$ & $-0.2 \pm 1.2 \pm 2.0$
\\
VR-Int & $18.7 \pm 0.9$ & $0.2 \pm 0.2$ & $-0.3 \pm 0.1 \pm 0.3$
\\
VR-High & $19.7 \pm 0.7$ & $0.04 \pm 0.04$ & $-0.2 \pm 0.2 \pm 0.5$
\\
VR-High-1J & $51.9 \pm 1.9$ & $0.003 \pm 0.002$ & $-1.9 \pm 0.2 \pm 0.4$
\\
SR-Low-b & $15.4 \pm 1.0$ & $0.001 \pm 0.000$ & $-0.08 \pm 0.05 \pm 0.06$
\\
SR-Int-a & $22.3 \pm 0.8$ & $0 \pm 0$ & $-0.4 \pm 0.2 \pm 0.6$
\\
SR-Int-b & $11.4 \pm 0.5$ & $0.002 \pm 0.002$ & $-0.3 \pm 0.1 \pm 0.3$
\\
SR-High-8-a & $2.4 \pm 0.2$ & $0 \pm 0$ & $0 \pm 0 \pm 0$
\\
SR-High-8-b & $2.4 \pm 0.3$ & $0 \pm 0$ & $0 \pm 0 \pm 0$
\\
SR-High-1J & $1.0 \pm 0.1$ & $0.004 \pm 0.003$ & $0 \pm 0 \pm 0$
\\
SR-$\llbb$ & $0.7 \pm 0.1$ & $0 \pm 0$ & $0 \pm 0 \pm 0$
\\
\end{tabular}
\caption[%
Background yields from prompt and fake/non-prompt sources in bins with
non-positive Matrix Method estimates
]{%
Background yields from prompt and fake/non-prompt sources in bins with
non-positive Matrix Method estimates.
The `MC FNP' yields are selected using truth information in simulated samples.
The `MM FNP' yields are from the Matrix Method, with statistical errors from
the sample weights and systematic variations from the larger absolute change
from variations of those weights.
}
\label{tab:2ljets_fnp_negative_mm}
\end{table}
